---
title: "Result Part 1: Downsampling Experiment"
output: html_notebook
---


# Downsampling Experiment

- datasets from Misharin, Zeisel, PBMC 10X
- Gene-Gene pearson Correlation
- Cell-Cell Spearman Correlation
- Stratified plots

##Read in Data 
```{r}
source('violin_plot.R')
source('my_functions.R')
library(colorout)
# Correlation matrix distance (takes a while to run)
ref.files <- list.files(path = 'Data',pattern = "ref", full.names = TRUE)
samp.files <- list.files(path = 'Data',pattern = "ds.rds", full.names = TRUE)
saver.files <- list.files(path = 'Data',pattern = "saver.rds", full.names = TRUE)
g2s3.files <- list.files(path = 'Data',pattern = "g2s3.rds", full.names = TRUE)
scimpute.files = list.files(path = 'Data',pattern = "scimpute.rds", full.names = TRUE)
viper.files <- list.files(path = "Data", pattern = "viper.rds", full.names = T)
dca.files <- list.files(path = "Data", pattern = "dca.rds", full.names = T)

# 10/10/19 add magic
magic.files = list.files(path = "Data", pattern = "magic.rds", full.names = T)

# 09/02/20 add scTSSR, saucie, knn_smooth, ALRA
sctssr.files = list.files(path = "Data", pattern = "sctssr", full.names = T)
saucie.files = list.files(path = "Data", pattern = "saucie.rds", full.names = T)
knnsmooth.files = list.files(path = "Data", pattern = "knnsmooth.rds", full.names = T)
alra.files = list.files(path = "Data", pattern = "alra", full.names = T)
enimpute.files = list.files(path = "Data", pattern = "enimpute", full.names = T)

datafiles = ls(pattern="\\.files")
for(d in datafiles){print(eval(parse(text = d)))}



dat <- vector("list", 3)
names(dat) <- c("Misharin", "PBMC", "Zeisel")


# Read in data and perform normalization
methods.name = c("saver", "g2s3", "scImpute", "viper", "dca", "magic", "sctssr", "saucie", "knnsmooth", "alra", "enimpute")


for (i in 1:3) {
  print(i)
  dat[[i]] <- vector("list", length(methods.name)*2+4)
  names(dat[[i]]) <- c("ref", "ds", methods.name, paste(c("ref", "ds", methods.name), "norm", sep = "_"))
  dat[[i]][[1]] <- readRDS(ref.files[i])
  dat[[i]][[2]] <- readRDS(samp.files[i])
  dat[[i]][[3]] <- readRDS(saver.files[i])$estimate
  dat[[i]][[4]] <-  readRDS(g2s3.files[i])
  dat[[i]][[5]] <- readRDS(scimpute.files[i])
  dat[[i]][[6]] <- readRDS(viper.files[[i]][[1]])
  dat[[i]][[7]] <- as.matrix(readRDS(dca.files[i])[-1,-1])
  dimnames(dat[[i]][[7]]) = dimnames(dat[[i]][[1]])
  dat[[i]][[8]] <- readRDS(magic.files[i])
  dat[[i]][[9]] <- readRDS(sctssr.files[i])
  dat[[i]][[10]] <- readRDS(saucie.files[i])
  dat[[i]][[11]] <- readRDS(knnsmooth.files[i])
  dat[[i]][[12]] <- readRDS(alra.files[i])

#----- use for first time calculation---------- (Slow!)
  # start_time <- Sys.time()
  # enimpute_datalist = dat[[i]][c(3,5,7,8,12)] #saver, scimpute,  dca, magic, alra
  # enimpute_res = Enimpute(enimpute_datalist)
  # end_time  <- Sys.time()
  # print(end_time - start_time)
  # saveRDS(enimpute_res$count.EnImpute.exp,file = gsub('_ds', '_ds_enimpute', samp.files[[i]]))
  # dat[[i]][[13]] <- enimpute_res$count.EnImpute.exp
#----------------------------------------------
  dat[[i]][[13]] <- readRDS(enimpute.files[i])
# #normalization
  for(j in 1:13){
    # print(dim(dat[[i]][[j]]))
    dat[[i]][[13+j]] <- normalizeData(dat[[i]][[j]])
  }
}



# remove small negative values from imputed data
for (i in 1:3){
  names(dat[[i]]) <- c("ref", "ds", methods.name, paste(c("ref", "ds", methods.name), "norm", sep = "_"))
  for (j in 1:length(dat[[i]])){
    temp = as.matrix(dat[[i]][[j]])
    temp[temp<0]=0
    dat[[i]][[j]]=temp
  }
}

# add gene and cellnames to the dataset
```

## Calculate Gene-wise and Cell-wise Correlation

```{r}

load('Data/dat4.rda')
cor.dat = vector("list", 2)
length(methods.name)

for (i in 1:2) {
  cor.dat[[i]] <- vector("list", 3)
  names(cor.dat[[i]]) <- names(dat)
  for (j in 1:3) {
    # compare 1 + 5methods
    cor.dat[[i]][[j]] <- vector("list", length(methods.name)+1)
    names(cor.dat[[i]][[j]]) <- c("obs", methods.name)
  }
}


for (i in 1:3) {
  for (j in 1:(length(methods.name )+1)) {
    # compare with normalized reference datset
      ind <- (length(methods.name)+4):length(dat[[i]])
      print(j)
      ref = dat[[i]][[length(methods.name)+3]]
      tmp = as.matrix(dat[[i]][[ind[j]]])
      if (sum(dim(ref)==dim(tmp))!=2){
      ref = ref[rownames(ref) %in% rownames(tmp), colnames(ref) %in% colnames(tmp)]}  #sctssr removed some genes
      cor.dat[[1]][[i]][[j]] <- get.cor.gene(ref, tmp, "pearson")
      cor.dat[[2]][[i]][[j]] <- get.cor.cell(ref, tmp, 'spearman')   
  }
}

str(cor.dat)
saveRDS(cor.dat, file = 'Data/cor_dat4.rds')

lapply(cor.dat[[2]], function(x) lapply(x, function(y) mean(y)))

lapply(cor.dat[[1]], function(x) lapply(x, function(y) mean(y)))



```


## Generate Plots: Figure 1

```{r}
cor.dat  <- readRDS("Data/cor_dat4.rds")
source("violin_plot.R")
source("my_functions.R")
library(tidyverse)
names(cor.dat[[1]][[1]])
n_methods = length(cor.dat[[1]][[1]])

# put g2s3 to the first comparison
# The order of tests: G2S3, SAVER, kNN-smoothing, MAGIC, scImpute, VIPER, ALRA, scTSSR, DCA, SAUCIE, EnImpute.
cor.dat= lapply(cor.dat , function(x) lapply(x, `[`, c(1, 3, 2, 10, 7, 4, 5, 11, 8, 6, 9, 12)))
names(cor.dat[[1]][[1]])



ds.plot = function(dim = "gene"){
  x <- c(1, 2, 3)
  plot(x, c(-5, -5, -5), type = "p", ylab = " ", xlab = " ", cex = 1.5, 
       ylim = c(0.0, 1), xlim = c(1, 9), lwd = 2, pch = 5, axes = FALSE, main = " ")

  axis(1, at = c(2.1, 4.6, 7.1), labels = FALSE)
  text(c(2.1, 4.6, 7.1), par()$usr[3]-0.03,
       labels = c("Reyfman", "PBMC", "Zeisel"), srt = 45, adj = 1, 
       xpd = TRUE, cex = 2)
  axis(2, pos = 1.1)
  par(las = 0)

  mtext("Correlation with Reference", side = 2, line = 1, cex = 2)


  fill <- c("white", "#cb181d", "#6baed6", '#8491B4B2', "#bae4b3", 'orange', "#FFD700", '#3399FF', "#CC79A7","#66CC99", '#D55E00', '#7570b3')
  xloc <- seq(1.5, by = 0.15, length.out = n_methods)

  idx = ifelse(dim=="gene", 1, 2)

  for (i in 1:3) {
    for (j in 1:n_methods) {
      boxplot.ej(cor.dat[[idx]][[i]][[j]], xloc = xloc[j]+2.2*(i-1), 
                 cex.boxpoint = ps, fill = fill[j],width.box = 0.1, width.hor = 0.1)
    }
  }

  yadj = 0.10+0.2
  xadj = 2.50
  l = seq(from = 0.4, by = -0.06, length.out = n_methods)
  r = l + 0.02
  t = l + 0.01
  imp_names = c('Observed',"G2S3", "SAVER", "kNN-smoothing", "MAGIC", "scImpute", "VIPER", "ALRA", "scTSSR", "DCA", "SAUCIE", "EnImpute")

  for (i in 1:n_methods){
    rect(5.25+xadj, l[i]+yadj, 5.35+xadj, r[i]+yadj, col = fill[i], lwd = 2)
    text(5.4+xadj, t[i]+yadj , imp_names[i], cex = 1.6, font = 1, adj = 0)
  }
}


pdf("Plots/FigureS1_rev.pdf",18, 14)
par(cex.main = 2, mar = c(7, 5, 2, 1) + 0.1, oma = c(1, 3, 2, 0), 
      mgp = c(3.5, 1, 0),
      cex.lab = 1.6, font.lab = 2, cex.axis = 1.6, mfrow = c(2,1))
ds.plot(dim = "gene")
title("Gene", line = -2, outer = TRUE)
ds.plot(dim = "cell")
title("Cell", line = -35, outer = TRUE)
dev.off()


#drop SAUCIE for figure 1 plot
cor.dat.v2 = cor.dat
for(i in 1:3){
  cor.dat.v2[[1]][[i]][[11]]=NULL
  cor.dat.v2[[2]][[i]][[11]]=NULL
}
n_methods = length(cor.dat.v2[[1]][[1]])
ds.plot2 = function(dim = "gene", cor.dat = cor.dat.v2){
  x <- c(1, 2, 3)
  plot(x, c(-5, -5, -5), type = "p", ylab = " ", xlab = " ", cex = 1.5, 
       ylim = c(0.0, 1), xlim = c(1, 9), lwd = 2, pch = 5, axes = FALSE, main = " ")

  axis(1, at = c(2.1, 4.6, 7.1), labels = FALSE)
  text(c(2.1, 4.6, 7.1), par()$usr[3]-0.03,
       labels = c("Reyfman", "PBMC", "Zeisel"), srt = 45, adj = 1, 
       xpd = TRUE, cex = 1.6)
  axis(2, pos = 1.1)
  par(las = 0)
  mtext("Correlation with Reference", side = 2, line = 1, cex = 1.6)

  fill <- c("white", "#cb181d", "#6baed6", '#8491B4B2', "#bae4b3", 'orange', "#FFD700", '#3399FF', "#CC79A7","#66CC99", '#7570b3')
  xloc <- seq(1.5, by = 0.15, length.out = n_methods)

  idx = ifelse(dim=="gene", 1, 2)

  for (i in 1:3) {
    for (j in 1:n_methods) {
      boxplot.ej(cor.dat[[idx]][[i]][[j]], xloc = xloc[j]+2.2*(i-1), 
                 cex.boxpoint = ps, fill = fill[j],width.box = 0.1, width.hor = 0.1)
    }
  }

  yadj = 0.10+0.2
  xadj = 2.50
  l = seq(from = 0.3, by = -0.06, length.out = n_methods)
  r = l + 0.02
  t = l + 0.01
  imp_names = c('Observed',"G2S3", "SAVER", "kNN-smoothing", "MAGIC", "scImpute", "VIPER", "ALRA", "scTSSR", "DCA", "EnImpute")

for (i in 1:n_methods){
    rect(5.25+xadj, l[i]+yadj, 5.35+xadj, r[i]+yadj, col = fill[i], lwd = 2)
    text(5.40+xadj, t[i]+yadj , imp_names[i], cex = 1.6, font = 1, adj = 0)
  }
}


pdf("Plots/Figure1_rev.pdf",18, 14)
par(cex.main = 2, mar = c(7, 5, 2, 1) + 0.1, oma = c(1, 3, 2, 0), 
      mgp = c(3.5, 1, 0),
      cex.lab = 1.6, font.lab = 2, cex.axis = 1.6, mfrow = c(2,1))
ds.plot2(dim = "gene", cor.dat.v2)
title("Gene", line = -2, outer = TRUE)
ds.plot2(dim = "cell", cor.dat.v2)
title("Cell", line = -35, outer = TRUE)
dev.off()


```

## Mean-Variance Relationship Preserved in Downsampling Data
```{r}
source('violin_plot.R')
source('my_functions.R')
library(colorout)
# Correlation matrix distance (takes a while to run)
ref.files <- list.files(path = 'Data',pattern = "ref", full.names = TRUE)
samp.files <- list.files(path = 'Data',pattern = "ds.rds", full.names = TRUE)

CV.corr = rep(0, length = 3)
CV = vector('list', 3)
names(CV.corr) = c("Reyfman", "PBMC", "Zeisel")
for (i in 1:3){
  r = readRDS(ref.files[[i]])
  d = readRDS(samp.files[[i]])
  r.cv = CV.all(r)
  d.cv = CV.all(d)
  CV.corr[i] = cor(r.cv, d.cv)  
  CV[[i]] = data.frame(ref = r.cv, samp = d.cv)
}
print(CV.corr)


library(ggpubr)
names(CV) <- c("Reyfman", "PBMC", "Zeisel")
cor_plots = lapply(seq_along(CV), function(i) {
  x = CV[[i]]
  ggplot(x, aes(x=ref, y=samp)) + 
    geom_point(color="gray36")+ xlim(0,5)+ ylim(0,5)+ xlab("CV (Before Down-sampling)") + ylab("CV (After Down-sampling)")+ ggtitle(names(CV)[i])+stat_cor(method = "pearson", label.x = 3, label.y = 1, aes(label = ..r.label..))+theme(plot.title = element_text(hjust = 0.5))
})


library(gridExtra)

pdf("Plots/Figure_S_Mean_Variance.pdf", width=9, height=3)
do.call("grid.arrange", c(cor_plots, ncol=3))
dev.off()

```



## Statified Correlation

```{r}
load('Data/dat4.rda')

# The order of tests: G2S3, SAVER, MAGIC, scImpute, VIPER, DCA

non_zero_prop = function(df){apply(df, 1, function(x)sum(x!=0, na.rm = T)/ncol(df))} #percent of expressed 


strate = vector("list", 3)
names(strate) = names(dat)

cutoff = matrix(c(.3, .8,
          .3, .8,
          .3, .8), 3,2, byrow = T)

for(i in 1:3){
  non_zero_rate = non_zero_prop(dat[[i]][[2]])
  # print(summary(non_zero_rate))
  id_low = which(non_zero_rate<=cutoff[i, 1])
  id_middle = which(non_zero_rate<=cutoff[i, 2]&non_zero_rate>cutoff[i, 1])
  id_high = which(non_zero_rate>cutoff[i, 2])
  strate[[i]][[1]] = id_low
  strate[[i]][[2]] = id_middle
  strate[[i]][[3]] = id_high
  cat("low: ", length(id_low), ", mid: ", length(id_middle), ", high", length(id_high), "\n" )
  names(strate[[i]]) <- c("low", "middle", "high")
}



nGene.strat = sapply(strate, function(x) sapply(x, length))
nGene.per = t(nGene.strat)/apply(nGene.strat, 2, sum)

# percentage of mildly + widely expressed genes
rownames(nGene.per)[1] = "Reyfman" 
apply(nGene.per[, 2:3], 1, sum)


cor.dat.strate <- vector("list", 2)
names(cor.dat.strate) <- c("gene", "cell")

for (i in 1:2) {
  cor.dat.strate[[i]] <- vector("list", 3)
  names(cor.dat.strate[[i]]) <- names(dat)
  for (j in 1:3) {
    cor.dat.strate[[i]][[j]] <- vector("list", length(dat[[i]])/2-1)
    names(cor.dat.strate[[i]][[j]])<- c("obs", methods.name)
    for (k in 1:length(cor.dat.strate[[i]][[j]])){
      cor.dat.strate[[i]][[j]][[k]]  <- vector('list', 3)
      names(cor.dat.strate[[i]][[j]][[k]]) = c('low', 'middle', 'high')
    }
  }
}

for (i in 1:3) {
  for (j in 1:(length(methods.name )+1)) {
    # compare with normalized reference datset
      ind <- (length(methods.name)+4):length(dat[[i]])
      print(j)
      ref = dat[[i]][[length(methods.name)+3]]
      tmp = as.matrix(dat[[i]][[ind[j]]])
      if (sum(dim(ref)==dim(tmp))!=2){
        ref = ref[rownames(ref) %in% rownames(tmp), colnames(ref) %in% colnames(tmp)]
      }  #sctssr removed some genes

      for(k in 1:3){
        if (ind[j]==22){ # deal with sctssr
          stratum = names(strate[[i]][[k]][names(strate[[i]][[k]])%in% rownames(ref)])}
        else{
          stratum = strate[[i]][[k]]
        }
      cor.dat.strate[[1]][[i]][[j]][[k]] <- get.cor.gene(ref[stratum,], tmp[stratum,], "pearson")
      cor.dat.strate[[2]][[i]][[j]][[k]] <- get.cor.cell(ref[stratum,], tmp[stratum,], 'spearman')   
    }
  }
}

  n_methods = length(cor.dat.strate[[1]][[1]])

cor.dat.strate= lapply(cor.dat.strate , function(x) lapply(x, `[`, c(1, 3, 2, 4: n_methods)))
names(cor.dat.strate[[1]][[1]])


saveRDS(cor.dat.strate, file = "Data/cor_dat_strate_final.rds")



```

**Functions for Generating Figure S1**
```{r}
corplot.strat = function(cor.dat, k){
  Gene_Levels = c("Rarely", "Mildly", "Widely")
  x <- c(1, 2, 3)
  par(las = 1)
  plot(x, c(-5, -5, -5), type = "p", ylab = " ", xlab = " ", cex = 1.5, 
       ylim = c(0.0, 1), xlim = c(1, 9), lwd = 2, pch = 5, axes = FALSE, main = " ")
   axis(1, at = c(2.3, 4.8, 7.3), labels = FALSE)
  text(c(2.1, 4.6, 7.1), par()$usr[3]-0.07,
       labels = c("Reyfman", "PBMC", "Zeisel"), srt = 45, adj = 1, 
       xpd = F, cex = 2)
  axis(2, pos = 1.1)

  n_methods = length(cor.dat.strate[[1]][[1]])

  fill <- c("white", "#cb181d", "#6baed6", '#8491B4B2', "#bae4b3", 'orange', "#FFD700", '#3399FF', "#CC79A7","#66CC99", '#D55E00', '#7570b3')
  xloc <- seq(1.5, by = 0.15, length.out = n_methods)


# Gene-to-Gene plot-------------------------------------------------------
  for (i in 1:3) {
    for (j in 1:n_methods) { #put G2S3 before SAVER
      boxplot.ej(cor.dat[[1]][[i]][[j]][[k]], xloc = xloc[j]+2.3*(i-1), 
                 cex.boxpoint = ps, fill = fill[j],width.box = 0.1, width.hor = 0.1)
    }
  }

  # configure legends
  yadj = 0.50
  xadj = 2.60
  l = seq(from = 0.26, by = -0.06, length.out = n_methods)
  r = l + 0.02
  t = l + 0.01
  imp_names = c('Observed',"G2S3", "SAVER", "kNN-smoothing", "MAGIC", "scImpute", "VIPER", "ALRA", "scTSSR", "DCA", "SAUCIE", "EnImpute")

# Draw Legend 
  if(k==1){
  for (i in 1:n_methods){
    rect(5.35+xadj, l[i]+yadj, 5.45+xadj, r[i]+yadj, col = fill[i], lwd = 2)
    text(5.5+xadj, t[i]+yadj, imp_names[i], cex = 1.6, font = 1, adj = 0)
  }
  }


  if(k==3){
  par(las = 0)
  mtext("Correlation with Reference",side = 2, line = 3, cex = 1.6)
  par(las = 1)}

  if(k==2){
    par(las = 1)
     mtext("Gene", side = 3, line = 1, cex = 2,font = 2)
  }

  x <- c(1, 2, 3)
  plot(x, c(-5, -5, -5), type = "p", ylab = " ", xlab = " ", cex = 2, 
       ylim = c(0.0, 1), xlim = c(1, 9), lwd = 2, pch = 5, axes = FALSE, main = " ")

  axis(1, at = c(2.3, 4.8, 7.3), labels = FALSE)
  text(c(2.1, 4.6, 7.1), par()$usr[3]-0.07,
       labels = c("Refyman", "PBMC", "Zeisel"), srt = 45, adj = 1, 
       xpd = TRUE, cex = 2)
  axis(2, pos = 1.1)

#  Cell-to-Cell plot-----------------------------------------------------

for (i in 1:3) {
    for (j in 1:n_methods) { #put G2S3 before SAVER
      boxplot.ej(cor.dat[[2]][[i]][[j]][[k]], xloc = xloc[j]+2.3*(i-1), 
                 cex.boxpoint = ps, fill = fill[j],width.box = 0.1, width.hor = 0.1)
    }
  }

  if(k==3){
    par(las = 0)
    # mtext("Gene (Pearson)",side = 2, line = 3, cex = 1.6, font = 2)
    mtext("Correlation with Reference",side = 2, line = 3, cex = 1.6)
    par(las = 1)
  }

  if(k==2){
    par(las = 1)
     mtext("Cell", side = 3, line = 0, cex = 2,font = 2)
  }

  par(las = 1)
  mtext(Gene_Levels[k], side = 1, line = 8.5, cex = 2,font = 2)

}

```


## Generate figure S1

```{r}
cor.dat.strate = readRDS("Data/cor_dat_strate_final.rds")
names(cor.dat.strate[[1]][[1]])
cor.dat.strate= lapply(cor.dat.strate , function(x) lapply(x, `[`, c(1, 2, 3, 10, 7, 4, 5, 11, 8, 6, 9, 12)))
names(cor.dat.strate[[1]][[1]])
pdf('Plots/Figure_S3_rev.pdf', 28, 12)
# pdf('Plots/Figure_S1_quantile.pdf', 16, 10)
par(mfcol = c(2, 3), cex.main = 2, mar = c(8, 5, 0, 0) + 0.1, oma = c(2, 0, 5, 0), 
    mgp = c(3.5, 1, 0),
    cex.lab = 2, font.lab = 2, cex.axis = 2, bty = "n", las = 1, xpd = FALSE)

for (k in 3:1){
  corplot.strat(cor.dat.strate, k)
  par(mar = c(8, 0, 0, 0))
}
par(las = 0)
dev.off()
```