---
title: "Result Part 4: Regulatory Network"
output: html_notebook
---


# Regulatory Netowrk Analysis

- datasets:
  + Paul et al. of Regulatory Genes
  + Buettner et al. of Cell Cycle Periodic Genes
- plot correlated/anticorrelated gene pairs in Paul dataset
- heatmap of all pairwise regulartory gene relationship
- fraction of buetttner periodic markers in correct direction


## Section 1: Paul Data Analysis

### Data Preparation

```{r}
load('Data/Paul_Cell_MARSseq_GSE72857.RData')
#find GATA1 gene
gata1 = 'gata1'

grep(gata1, info.genes, value = T, ignore.case = T)

#find PU.1 gene
pu.1 <- c('SFPI1', 'SPI-1', 'SPI-A', 'PU.1', 'OFF')
lapply(pu.1, function(x) grep(x, info.genes, value = T, ignore.case = T))
#double check existence
antigenes <- c("Gata1", "Sfpi1")
lapply(antigenes, function(x)grep(x, info.genes, value = T, ignore.case = T))

gene.names <-sapply(strsplit(rownames(data.debatched), ";"), "[", 1)
is.informative <- gene.names %in% info.genes[order(info.genes)]
data.info.genes <- data.debatched[is.informative,]
rownames(data.info.genes) <- gene.names[is.informative]

dim(data.info.genes)
quantile.cell <- quantile(colSums(data), 0.2)
quantile.cell

write.table(data.info.genes, file = 'Data/Paul_infogene.txt', row.names = F, col.names = F, quote = F)
saveRDS(data.info.genes, file = 'Data/paul.rds')

```

### Read in Data

```{r}
#readin Data
paul <- readRDS('Data/paul.rds')
paul_g2s3 <- read.csv('Data/Paul_g2s3.csv', header = F)
paul_scimpute = readRDS('Data/Paul_scimpute.rds')
paul_dca <- read.csv('Data/paul_dca.tsv', sep = '\t')
paul_magic <- read.csv('Data/Paul_magic.csv');paul_magic = t(paul_magic[,-1])
dimnames(paul_g2s3) <- dimnames(paul)
dimnames(paul_scimpute) <- dimnames(paul)
dimnames(paul_dca) <- dimnames(paul)
dimnames(paul_magic)  <- dimnames(paul); 
# paul_dca = as.matrix(paul_dca)
paul = data.frame(t(paul))
paul_g2s3 = data.frame(t(paul_g2s3))
paul_scimpute = data.frame(t(paul_scimpute))
paul_dca = data.frame(t(paul_dca))
paul_magic = data.frame(t(paul_magic))
save(paul, paul_g2s3, paul_scimpute, paul_dca, paul_magic, file = "Data/paul_allMethod.rda")
```

### Generate Figure 6

*revision: add 5 new methods*
```{r}
# USE R-3.4.2
load('Data/paul_allMethod.rda')
library(tidyverse)
# sctssr <- readRDS("Data/paul_sctssr.rds") %>% t()
# saucie <- read.csv("Data/paul_saucie.csv", row.names = 1)
# saveRDS(saucie, file = "Data/paul_saucie.rds")
saucie <- readRDS("Data/paul_saucie.rds") %>% t()
saver <- readRDS("Data/Paul_saver.rds")$estimate  %>% t() 
knnsmooth <- readRDS("Data/paul_knnsmooth.rds")  %>% t()
alra <- readRDS("Data/paul_alra.rds") %>% t()
# sctssr removed some gene, use original data to show
tmp <- readRDS("Data/paul_sctssr.rds")%>%t()
sctssr = paul; 
sctssr[, c(colnames(sctssr) %in% colnames(tmp))] %>%dim()
sctssr[, c(colnames(sctssr) %in% colnames(tmp))] = tmp
sctssr <- sctssr 
viper <- readRDS("Data/paul_viper.rds") %>%t()
saucie <- readRDS("Data/paul_saucie.rds")%>%t()

enimpute <- readRDS("Data/paul_enimpute.rds")
dimnames(enimpute) <- dimnames(saver)


# order of methods:  G2S3, SAVER, kNN-smoothing, MAGIC, scImpute, VIPER, ALRA, scTSSR, DCA, SAUCIE, EnImpute
paul_list <- list(paul, paul_g2s3, saver, knnsmooth, paul_magic, paul_scimpute, viper, alra, sctssr, paul_dca, saucie, enimpute)
names(paul_list) <- c("Raw", "G2S3", "SAVER", "kNN-smoothing", "MAGIC", "scImpute", "VIPER", "ALRA", "scTSSR", "DCA", "SAUCIE", "EnImpute")

featureGenes = c('Gfi1', 'Fli1', "Cebpa", 'Sfpi1', 'Gata2', 'Egr1', 'Tal1', 'Zfpm1', 'Gata1', 'Klf1')
library(ggplot2)
library(gridExtra)

heatplot = function(df, title){
  library(reshape2)
  df_features = df[,featureGenes]
  cor.raw = cor(df_features)
  melted_cor.raw  <- melt(cor.raw )
  p = ggplot(data = melted_cor.raw , aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
     midpoint = 0, limit = c(-1,1), space = "Lab", 
     name="Pearson\nCorrelation", breaks = c(-1, 0, 1))+
     theme_bw(base_size = 20) + theme(axis.title.x = element_blank(),
            axis.title.y = element_blank()) + ggtitle(title) +theme(plot.title = element_text(hjust = 0.5, size = 24, face = "bold")) + theme(legend.title = element_text(size = 16),axis.text.x = element_text(angle = 90, hjust = 1)) + theme(plot.margin = unit(c(0.3,0.3,1,0.3), "cm"))
  p = p + theme(legend.position = "top")
  return(p)
}

heatplots <- lapply(seq_along(paul_list), function(i) {print(i);heatplot(paul_list[[i]], names(paul_list)[i])})

pdf('Plots/Figure6_rev2.pdf',width=30, height=15)
do.call("grid.arrange", c(heatplots, ncol=6))
dev.off()
```

### Generate Figure S5 and S6
*(Anti and Cor pair genes, all)*
```{r}

library(ggpubr)

# Figure 6 appendix

antipair = list(c('Fli1', 'Klf1'), c('Egr1', 'Gfi1'), c('Cebpa','Gata1'), c('Sfpi1','Gata1' ))
copair = list(c('Sfpi1', "Cebpa"), c("Zfpm1", 'Gata1'), c('Klf1','Gata1' ))
antipairPlot = function(df, title, antigenes, col){
  ce <- cor(df[, antigenes[1]], df[, antigenes[2]])
  p = qplot(df[, antigenes[1]], df[, antigenes[2]], xlab = antigenes[1], ylab = antigenes[2], alpha=I(0.4), colour = I(col)) + theme_bw(base_size = 24)+
  ggtitle(paste0(title, '\n', "Cor: ", round(ce,2))) +theme(plot.title = element_text(hjust = 0.5))+theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold")) + theme(plot.margin = margin(0.5, 2, 1, 2, "cm")) + theme(axis.title.x = element_text(face = 'bold'), axis.title.y = element_text(face = 'bold'), axis.text.y=element_blank(),axis.text.x=element_blank())   + theme(plot.margin = unit(c(0.3,0.5,1,1), "cm"))
  return(p)
}


pairPlot = function(df, title, cogenes, col){
  ce <- cor(df[, antigenes[1]], df[, antigenes[2]])
  p = qplot(df[, cogenes[1]], df[, cogenes[2]], xlab = cogene[1], ylab = cogene[2],alpha=I(0.4), colour = I(col)) + theme_bw(base_size = 24)+
  ggtitle(paste0(title, '\n', "Cor: ", round(ce,2))) +theme(plot.title = element_text(hjust = 0.5))+theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold")) + theme(plot.margin = margin(0.5, 2, 1, 2, "cm")) + theme(axis.title.x = element_text(face = 'bold'), axis.title.y = element_text(face = 'bold'), axis.text.y=element_blank(),axis.text.x=element_blank()) + theme(plot.margin = unit(c(0.3,0.5,1,1), "cm"))
  return(p)
}

mycolors = ggsci::pal_lancet()(9)

antiplots_1  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], antipair[[1]], mycolors[1]))
antiplots_2  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], antipair[[2]], mycolors[2]))
antiplots_3  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], antipair[[3]],mycolors[3]))
antiplots_4  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], antipair[[4]],mycolors[4]))

coplots_1  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], copair[[1]],mycolors[5]))
coplots_2  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], copair[[2]],mycolors[6]))
coplots_3  <- lapply(seq_along(paul_list), function(i) antipairPlot(paul_list[[i]], names(paul_list)[i], copair[[3]],mycolors[7]))

library(grid)
pdf('Plots/FigureS5_antipair.pdf',width=40, height=15)
do.call("grid.arrange", c(c(antiplots_1, antiplots_2, antiplots_3, antiplots_4), ncol=12))
dev.off()

pdf('Plots/FigureS6_copair.pdf',width=40, height=10)
do.call("grid.arrange", c(c(coplots_1, coplots_2, coplots_3), ncol=12))
dev.off()

```


## Section 2: Buettner Data Analysis
### Data preparation

```{r}
# read in gene annotation information
library("readxl")
library(tidyverse)
library(janitor)
markers  <- read_excel('Data/41422_2016_BFcr201684_MOESM26_ESM.xlsx', skip = 4)%>% select(1:5)%>%clean_names() %>% filter(core_67=="YES") %>% data.frame()
head(markers)
table(markers$stage)
saveRDS(markers, file = "Data/core67.rds")

# read in Buettner data
bd = list.files("../../Raw_data/Bruettner/E-MTAB-2805.processed.1", full.names = T)
bd.names  <- list.files("../../Raw_data/Bruettner/E-MTAB-2805.processed.1")

bd.dat  <- list('vector', 3)
for(i in 1:length(bd)){
  bd.dat[[i]] <- read.table(bd[i], header = T)%>%mutate(AssociatedGeneName = str_to_upper(AssociatedGeneName))%>%data.frame()
}
names(bd.dat) <- sapply(strsplit(bd.names, "_"), '[', 1)

head(bd.dat[[1]])

sum(markers$gene_name %in% bd.dat[[1]]$AssociatedGeneName)
lapply(bd.dat, dim)

saveRDS(bd.dat, file = "Data/Buettner.rds")

b  <- readRDS("Data/Buettner.rds")
b.i  <- lapply(b, function(x) {
  # remove unannotated genes
  idx = which(is.na(x$AssociatedGeneName))
  x = x[-idx,]
  res <- x[,5:ncol(x)]
  udx <- !duplicated(x$AssociatedGeneName)
  res <- res[udx,]; rownames(res) <- x[udx,"AssociatedGeneName"]
  return(res)}
  )
bue <- do.call(cbind, b.i) %>% data.frame

sum(markers$gene_name %in% rownames(bue))


bue[1:5, 1:5]
#basic filtering
summarize <- function(df){
  zero_prop = apply(df, 1, function(x)sum(x==0, na.rm = T))/ncol(df)
  cat("* summary of zero proportion\n")
  print(summary(zero_prop))
  lib_size = apply(df,2, function(x)sum(x, na.rm = T))
  cat("* summary of library size\n")
  print(summary(lib_size))
  return(list(zero_prop = zero_prop,lib_size = lib_size))
}
summary.list = summarize(as.matrix(bue))
gene.fil <- summary.list$zero_prop<0.80
bue <- bue[gene.fil,]

sum(markers$gene_name %in% rownames(bue))
saveRDS(bue, file = "Data/bue_dat.rds")
write.csv(bue, file = "Data/bue_dat.csv")
write.table(bue, file = "Data/bue_dat.tsv", sep = "\t")

```

### Generate Table 2
```{r}

raw <- readRDS("Data/bue_dat.rds")
knnsmooth <- readRDS("Data/bue_dat_knnsmooth.rds")
magic <- readRDS("Data/bue_dat_magic.rds")
dca <- readRDS("Data/bue_dat_dca.rds")
g2s3 <- readRDS("Data/bue_dat_g2s3.rds")
saver <- readRDS("Data/bue_dat_saver.rds")$estimate
scimpute <- readRDS("Data/bue_dat_scimpute.rds")
alra <- readRDS("Data/bue_dat_alra.rds")
sctssr <- readRDS("Data/bue_dat_sctssr.rds")
saucie <- readRDS("Data/bue_dat_saucie.rds")
viper <- readRDS("Data/bue_dat_viper.rds")
enimpute <- readRDS("Data/bue_dat_enimpute.rds")

info <- readRDS("Data/core67.rds")
table(info$stage)
st <- c('G1-S', 'G2-M')
info <- info[which(info$gene_name %in% rownames(raw)),]


calc.cor <- function(x, p = p){
  x[x<0]=0
  x = log10(x+0.001)
  cellstage = grep(p, colnames(x))

  x.1 <- x[match(info[which(info$stage=="G1-S"), 'gene_name'], rownames(x)),cellstage]
  x.2 <- x[match(info[which(info$stage=="G2-M"), 'gene_name'], rownames(x)),cellstage]

  if (p == "G1|S"){
    cors = cor(t(x.1), t(x.1))
  }else{
    cors = cor(t(x.2), t(x.2))
  }
  cors = c(cors[lower.tri(cors)])
  pos.cor <- median(cors)
  pos.prop <- sum(cors > 0)/length(cors)

  anti.cors <- cor(t(x.1), t(x.2))
  anti.cors  <- c(anti.cors[lower.tri(anti.cors)])
  neg.cor = median(anti.cors)
  neg.prop <- sum(anti.cors < 0)/length(anti.cors)

  return(list(pos.prop= pos.prop, neg.prop = neg.prop))
}


calc.cor2 <- function(x){
  x[x<0]=0
  x = log10(x+0.001)

  x.1 <- x[match(info[which(info$stage=="G1-S"), 'gene_name'], rownames(x)),]
  x.2 <- x[match(info[which(info$stage=="G2-M"), 'gene_name'], rownames(x)),]

  cors1 = cor(t(x.1), t(x.1))
  cors2 = cor(t(x.2), t(x.2))

  cors1 = c(cors1[lower.tri(cors1)])
  cors2 = c(cors2[lower.tri(cors2)])
  cors = c(cors1, cors2)
  pos.prop <- sum(cors > 0)/length(cors)

  anti.cors <- cor(t(x.1), t(x.2))
  anti.cors  <- c(anti.cors)
  neg.prop <- sum(anti.cors < 0)/length(anti.cors)
  return(list(pos.prop= pos.prop, neg.prop = neg.prop))
}

data.list = list(raw = raw, g2s3 = g2s3,saver  =saver, knnsmooth = knnsmooth, magic = magic,scimpute = scimpute, viper = viper, alra = alra, sctssr = sctssr, dca = dca, saucie = saucie, enimpute = enimpute)
names(data.list) <- c("Raw", "G2S3", "SAVER", "kNN-smoothing", "MAGIC", "scImpute", "VIPER", "ALRA", "scTSSR", "DCA", "SAUCIE", "EnImpute")

library(plyr)

res.list = lapply(data.list, function(x) calc.cor(x,"G1|S"))
dfs <- lapply(res.list , data.frame, stringsAsFactors = FALSE)
library(plyr)
r1<- rbind.fill(dfs); rownames(r1) <- names(data.list)
r1

cellstage <- grep("G2M", colnames(raw))
res.list = lapply(data.list, function(x) calc.cor(x,"G2M" ))
dfs <- lapply(res.list , data.frame, stringsAsFactors = FALSE)
r2<- rbind.fill(dfs); rownames(r2) <- names(data.list)
r2

res.list <- lapply(data.list, function(x) calc.cor2(x))
dfs <- lapply(res.list , data.frame, stringsAsFactors = FALSE)
r<- rbind.fill(dfs); rownames(r) <- names(data.list)
r

res.table = cbind(r1, r2)
rownames(res.table)  <- names(res.list)
write.csv(r, file = "Bue_results_rev.csv")

```
